---
title: "Hawaiian SFPW movement analysis"
author: "Amy Van Cise"
date: "April 18, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
  This workshop is an introduction to the use of the crawl package to 


```{r read data, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
rm(list=ls())
library(tidyverse)
library(lubridate)
library(magrittr)
library(stringr)

setwd("C:/Users/Amy/Documents/1 SIO/6 Year/crawl workhop")
data<-readr::read_csv("GmAll-Tag167-Complete_r15d3lc2.csv")
colnames(data)<-tolower(colnames(data))

##specify date formate using lubridate
data<-data %>%
  dplyr::mutate(date2=paste(data$date,data$time, sep=" ")) %>%
  dplyr::mutate(datetime=lubridate::mdy_hms(date2))

#count number of tags
ntags<- data$animal %>% unique() %>% length()

#convert location quality to 3,2,1,A,B
data$lc94<-data$lc94 %>% 
  replace(list=which(data$lc94=="DP"),values="L3") %>% 
  str_sub(start=-1)

# data<-cbind.data.frame(data$animal, data$datetime, data$latitude, data$longitud, data$lc94)
# colnames(data)<-c("animal","datetime","latitude","longitud","lc94")

##convert dataframe into a spatial data class (sf)
spdata<-data %>% sf::st_as_sf(coords=c("longitud","latitude")) %>% 
  sf::st_set_crs(.,4326)
  


```
###Data included
  Tag data from short-finned pilot whales throughout the Main Hawaiian Isalnds. There are `r ntags` tags included in this dataset. 

## Plot of data

```{r pressure, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
library(leaflet)
library(sf)
library(sp)
library(mapview)
library(ggthemes)


pal <- colorFactor(ggthemes::ptol_pal()(2), 
                   domain = spdata$animal)

m<-spdata %>% 
  #group_by(animal) %>% 
  #nest() %>% 
  #sample_n(1) %>% 
  #unnest() %>% 
  leaflet() %>%
  addProviderTiles(provider="Esri.OceanBasemap") %>% 
  addCircleMarkers(radius=1,weight=2, opacity=0.5, color = 'red') %>%
  #addPolylines(lng=~x, lat=~y, group = ~animal, color='red') %>% 
  addLegend(pal = pal,values = ~animal,labels = ~animal)

m

```

###Crawl model fit
```{r crawlinput, warning=FALSE, message=FALSE, echo=FALSE}
library(crawl)
library(pander)

spdata<-spdata %>% sf::st_transform(2785)

##function to add x and y columns to spatial data class
sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- do.call(rbind,sf::st_geometry(x))
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == ncol(ret))
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}

spdata<-spdata %>% sfc_as_cols

#nest spdata
spdata<-spdata %>% 
  group_by(animal) %>% 
  nest()


## set initial parameters for model
init_params <- function(d) {
  ret <- list(a = c(d$x[1], 0,
                    d$y[1], 0),
              P = diag(c(10 ^ 2, 10 ^ 2,
                         10 ^ 2, 10 ^ 2)))
  ret
} 

##attach parameters to the dataset

spdata <- spdata %>% 
  dplyr::mutate(init = purrr::map(data,init_params)
                )

#make sure errors are correctly ordered
order_lc<-function(x) {
  x$lc94 = factor(x$lc94, levels = c("3","2","1","0","A","B"))
  return(x)
}

spdata<-spdata %>% 
  dplyr::mutate(data=purrr::map(data,order_lc))

#set priors for all the parameters to be estimated

fit_crawl <- function(d, init) {

prior<-function(p){
  dnorm(p[1],log(250), 0.2, log=TRUE) +
  dnorm(p[2],log(500), 0.2, log=TRUE) +
  dnorm(p[3],log(1500), 0.2, log=TRUE) +
  dnorm(p[4],log(2500), 0.4, log=TRUE) +
  dnorm(p[5],log(2500), 0.4, log=TRUE) +
  dnorm(p[6],log(2500), 0.4, log=TRUE) +
  dnorm(p[8],-4, 2, log=TRUE)
}
    
fit<- crawl::crwMLE(
  mov.model = ~1,
  err.model = list(x= ~lc94-1),
  data=d,
  method="Nelder-Mead",
  Time.name = "datetime",
  initial.state = init,
  prior = prior,
  attempts = 8,
  control = list(
    trace = 0
    ),
  initialSANN = list(
    maxit = 1500,
    trace = 0
    )
)        
  fit
  }

spdatafit <- spdata %>% 
  sample_n(5) %>% 
  dplyr::mutate(fit = purrr::map2(data,init, fit_crawl),
                params = map(fit, crawl::tidy_crwFit))

##table of parameters
panderOptions('knitr.auto.asis', FALSE)
spdatafit$params %>% 
  walk(pander::pander,caption = "crwMLE fit parameters")

```

###Use the fit parameters to predict paths for two animals

```{r crawlpredictions, warning=FALSE, message=FALSE, echo=FALSE}

##crawl predictions
spdatafit <- spdatafit %>% 
  dplyr::mutate(predict = purrr::map(fit,
                                     crawl::crwPredict,
                                     predTime = '1 hour')) 

#graphs of x and y vs time
spdatafit$predict %>% purrr::walk(crawl::crwPredictPlot)

##function to change predictions to a spatial object
as.sf <- function(p,id,epsg,type,loctype) {
  p <- 
    sf::st_as_sf(p, coords = c("mu.x","mu.y")) %>% 
    dplyr::mutate(TimeNum = lubridate::as_datetime(TimeNum),
                  deployid = id) %>% 
    dplyr::rename(pred_dt = TimeNum) %>% 
    filter(loctype %in% loctype) %>% 
    sf::st_set_crs(.,epsg)
  if (type == "POINT") return(p)
  if (type == "LINE") {
    p <- p %>% dplyr::arrange(pred_dt) %>%  
      sf::st_geometry() %>% 
      st_cast("MULTIPOINT",ids = as.integer(as.factor(p$deployid))) %>% 
      st_cast("MULTILINESTRING") %>% 
      st_sf(deployid = unique(p$deployid))
    return(p)
  }
}

#convert fit to spatial object using above function
spdatafit <- spdatafit %>% 
  dplyr::mutate(sf_points = purrr::map2(predict, animal,
                                       as.sf,
                                       epsg = 2785, 
                                       type = "POINT",
                                       loctype = "p"),
                sf_line = purrr::map2(predict, animal,
                                     as.sf,
                                     epsg = 2785,
                                     type = "LINE",
                                     loctype = "p"))

#pull predicted lines from spdata and rbind the lines together
sf_pred_lines <- spdatafit$sf_line %>% 
  lift(rbind)() %>% 
  sf::st_set_crs(2785) 

n <- length(unique(sf_pred_lines$deployid)) #calculate # of individuals
pal <- colorFactor(ggthemes::ptol_pal()(n), 
                   domain = sf_pred_lines$deployid) #set color pallete

#map of paths
m <- sf_pred_lines %>% 
  sf::st_transform(4326) %>% 
  leaflet() %>% 
  addProviderTiles("Esri.OceanBasemap") %>% 
  addPolylines(weight = 2, color = ~pal(deployid)) %>% 
  addLegend(pal = pal,values = ~deployid,labels = ~deployid) 
  #suspendScroll()

m

```

##static map of the data
```{r crawlmap static, warning=FALSE, message=FALSE, echo=FALSE}
library(nPacMaps)
library(raster)
library(gdistance)
```

##create and map simulated tracks from error distribution using crawlr
```{r crawlr sim tracks, warning=FALSE, message=FALSE, echo=FALSE}
library(crawlr)

getsims<-function(x, iter, fixpath=FALSE, basemap) {
  predTimes <- seq(
    lubridate::ceiling_date(min(x$data$datetime), "hour"),
    lubridate::floor_date(max(x$data$datetime), "hour"),
    "1 hour"
    )
trk<-crawlr::get_sim_tracks(x, iter=iter, predTimes)
}

tracks<-getsims(spdatafit$fit[[1]], 20)

sim_points <- crawlr::get_sim_points(tracks, locType = "p", CRS("+init=epsg:2785"))

##map of predicted lines

#convert to sf object
sim_points <- sf::st_as_sf(sim_points)
sim_points <- sf::st_transform(sim_points,4326)

#build map using leaflet
m<-sf_pred_lines %>% 
  sf::st_transform(4326) %>% 
  leaflet() %>% 
  addProviderTiles("Esri.OceanBasemap") %>% 
  addCircleMarkers(data = sample_frac(sim_points, 0.25), radius = 2, weight = 2, 
                   opacity = 3, stroke = FALSE,
                   color = "purple") %>% 
  addPolylines(weight = 2, color = ~pal(deployid)) %>% 
  addLegend(pal = pal, values = ~deployid, labels = ~deployid)

m
  
```

